### 2021-06-17 paper_p.Rmd
First commit CSS files and added paper_p.

---
title: "Writing a reproducible paper in Pagedown^[Corresponding address: mail@paulcbauer.eu.<br> Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.]"
shorttitle: "**foo**:  A Capitalized Title"
author:
  - name: Paul C. Bauer
    affiliation: Mannheim Centre for European Social Research
    email: <mail@paulcbauer.eu>
  - name: Camille Landesvatter
    affiliation: Mannheim Centre for European Social Research
    email: <landesvatterc@gmail.com>
abstract: \noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below I outline some of the "tricks"/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document. The underlying files which produce this document can be downloaded [here]<https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).\vspace{.8cm}
# at least one keyword must be supplied
keywords: [keywords, not capitalized, "[Java]{.proglang}"]
paged-footnotes: true
link-citations: true
linestretch: 1.2
colorlinks: true
bibliography: index.bib
output:
  pagedown::html_paged:
    toc: true
    toc_depth: 3    # change to true for a self-contained document, but it'll be a litte slower for Pandoc to render
    self_contained: TRUE
    css: ['default.css', 'default-fonts.css', 'default-page.css']
# uncomment this line to produce HTML and PDF in RStudio:
#knit: pagedown::chrome_print
---

# Introduction

This template demonstrates some of the basic [markdown]{.proglang} you'll need to know to create a JSS article with **pagedown** [@R-pagedown].

## Code formatting

You should use **Pandoc**'s [markdown]{.proglang} but you also can use $\LaTeX$ commands:

* [Java]{.proglang} or $\proglang{Java}$ 
* **plyr**, [plyr]{.pkg} or $\pkg{plyr}$
* `print("abc")` or $\code{print("abc")}$

# [R]{.proglang} code

Can be inserted in regular [R Markdown]{.proglang} blocks.

```{r}
x <- 1:10
x
```

# Render the document

If you use the option `self_contained: false` (see line \#22 of this `Rmd` file), don't click on the `Knit` button in RStudio. Use instead the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.

# References {-}

```{r, include=FALSE}
knitr::write_bib(c(.packages(), 'pagedown', 'xaringan'), 'index.bib')
```